---
title: "SpatialDE on Mouse Olfactory Bulb Data (SingleCellExperiment)"
author: "Lambda Moses"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{SpatialDE on Mouse Olfactory Bulb Data (SingleCellExperiment)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For an introduction to what `SpatialDE` is and how to install it, see the [vignette on running SpatialDE on matrices](mouseob.html). In this vignette, we run `SpatialDE` on a `SingleCellExperiment` object.

## Normalize data
```{r, message=FALSE}
library(SpatialDE)
library(dplyr)
library(ggplot2)
library(reticulate)
library(Linnorm)
library(zeallot)
library(SingleCellExperiment)
theme_set(theme_bw())
```

```{r}
# Load the data
# Gene counts in each cell
data("mouseob")
# Locations of the cells
data("locations")
str(mouseob)
str(locations)
```

```{r}
# Create SingleCellExperiment object
sce <- SingleCellExperiment(assays = list(counts = t(mouseob)))
# Add locations to column metadata
colData(sce) <- DataFrame(locations)
```

```{r}
# Remove genes with low expression
gene_tot_counts <- rowSums(counts(sce) > 0)
sce <- sce[gene_tot_counts >= 3,]
```
How many genes are left after filtering?
```{r}
sce
```

```{r}
# Normalize data
normcounts(sce) <- Linnorm(counts(sce))
```

## Run `SpatialDE`
```{r, results='hide'}
# Add log total counts to colData
colData(sce)$log_total_counts <- log(colData(sce)$total_counts)
results <- RunSpatialDE(sce, covariates = "log_total_counts")
```

```{r}
head(results)
```

What are the most significant genes?
```{r}
results %>% 
  arrange(qval) %>% 
  select(g, l, qval) %>% 
  head()
```

```{r}
top_de <- results$g[order(results$qval)][1:4]
# Plot the most significant genes
MultiGenePlot(sce, genes_plot = top_de)
```

Contrast this to non-spatialDE genes
```{r}
bottom_de <- results$g[order(-results$qval)][1:4]
# Plot the most significant genes
MultiGenePlot(sce, bottom_de)
```

## Model search
```{r}
# Get significant results from SpatialDE
sig_res <- results %>% 
  filter(qval < 0.05)
```

```{r, results='hide'}
# Also try the linear and periodic models
ms_results <- ModelSearch(sce, sig_res, covariates = "log_total_counts")
```

```{r}
ms_results
```

```{r}
FSV_sig(results, do_label = TRUE, ms_results = ms_results, covariate_names = "log_total_counts")
```

There're 2 genes for which the periodic model fit better than the general model
```{r, fig.height=3}
periodic <- ms_results$g[ms_results$model == "PER"]
MultiGenePlot(sce, periodic)
```

## Automatic expression histology
What length scales (`l`) are present?
```{r}
sig_res %>% 
  group_by(l) %>% 
  dplyr::count()
```

```{r}
mean(sig_res$l)
```

```{r}
py_set_seed(456)
c(histology_results, patterns) %<-% AEHSpatialPatterns(sce, sig_res, covariates = "log_total_counts",
                                                       C = 3L, l = 1.5, verbosity = 1L)
```

```{r}
histology_results %>% 
  arrange(pattern)
```

```{r}
# Select top gene from each pattern to plot
(plot_genes <- histology_results %>% 
  left_join(sig_res, by = "g") %>% 
  group_by(pattern) %>% 
  summarize(g = g[which.min(qval)]))
```

```{r}
MultiGenePlot(sce, genes_plot = plot_genes$g, facet_titles = paste0("pattern", plot_genes$pattern))
```

```{r}
sessionInfo()
```
