---
title: "SpatialDE on Mouse Olfactory Bulb Data (Seurat)"
author: "Lambda Moses"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SpatialDE on Mouse Olfactory Bulb Data (Seurat)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For an introduction to what `SpatialDE` is and how to install it, see the [vignette on running SpatialDE on matrices](mouseob.html). In this vignette, we run `SpatialDE` on a `Seurat` object. Note that while Seurat 3.0 is to be released soon, with many really cool features, we are using Seurat 2.3.4 -- the CRAN version -- here, because Bioconductor requires that all dependencies must be from CRAN or Bioconductor. I'll update this vignette once Seurat 3.0 gets to CRAN.

## Normalize data
```{r, message=FALSE}
library(SpatialDE)
library(dplyr)
library(ggplot2)
library(reticulate)
library(Linnorm)
library(zeallot)
library(Seurat)
theme_set(theme_bw())
```

```{r}
py_config()
```

```{r}
# Load the data
# Gene counts in each cell
data("mouseob")
# Locations of the cells
data("locations")
str(mouseob)
str(locations)
```

We can use `NormalizeData` from `Seurat` to do the classical `log1p` normalization, we're using a more sophisticated data normalization method called `Linnorm` here. `SpatialDE` will use the normalized data in slot `data` rather than scaled data.
```{r}
# Create Seurat object
seu <- CreateSeuratObject(t(mouseob), min.cells = 3)
seu@data <- Linnorm(seu@raw.data)
```

## Run `SpatialDE`
```{r}
# Add log total counts to metadata
locations <- locations %>% 
  mutate(log_total_counts = log(total_counts))
rownames(locations) <- locations$X1
seu <- AddMetaData(seu, locations)
```

```{r, results='hide'}
results <- RunSpatialDE(seu, covariates = "log_total_counts")
```

```{r}
head(results)
```

What are the most significant genes?
```{r}
results %>% 
  arrange(qval) %>% 
  select(g, l, qval) %>% 
  head()
```

```{r}
top_de <- results$g[order(results$qval)][1:4]
# Plot the most significant genes
MultiGenePlot(seu, genes_plot = top_de)
```

Contrast this to non-spatialDE genes
```{r}
bottom_de <- results$g[order(-results$qval)][1:4]
# Plot the most significant genes
MultiGenePlot(seu, bottom_de)
```

## Model search
```{r}
# Get significant results from SpatialDE
sig_res <- results %>% 
  filter(qval < 0.05)
```

```{r, results='hide'}
# Also try the linear and periodic models
ms_results <- ModelSearch(seu, sig_res, covariates = "log_total_counts")
```

```{r}
ms_results
```

```{r}
FSV_sig(results, do_label = TRUE, ms_results = ms_results, covariate_names = "log_total_counts")
```

There're 2 genes for which the periodic model fit better than the general model
```{r, fig.height=3}
periodic <- ms_results$g[ms_results$model == "PER"]
MultiGenePlot(seu, periodic)
```

## Automatic expression histology
What length scales (`l`) are present?
```{r}
sig_res %>% 
  group_by(l) %>% 
  dplyr::count()
```

```{r}
mean(sig_res$l)
```

```{r}
py_set_seed(456)
c(histology_results, patterns) %<-% AEHSpatialPatterns(seu, sig_res, covariates = "log_total_counts",
                                                       C = 3L, l = 1.5, verbosity = 1L)
```

```{r}
histology_results %>% 
  arrange(pattern)
```

```{r}
# Select top gene from each pattern to plot
(plot_genes <- histology_results %>% 
  left_join(sig_res, by = "g") %>% 
  group_by(pattern) %>% 
  summarize(g = g[which.min(qval)]))
```

```{r}
MultiGenePlot(seu, genes_plot = plot_genes$g, facet_titles = paste0("pattern", plot_genes$pattern))
```

```{r}
sessionInfo()
```
